---
title: "320 Final Project"
author: "Amy, Katie, Shawdi"
date: "5/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```

## R Markdown


```{r message=FALSE, warning=FALSE}
library(rvest)
library(tidyverse)
url <-"html_top100.txt"
college_urls <- url %>%
  read_html() %>%
  html_node("body") %>% html_nodes("ol[class~=bEyEue]") %>% html_nodes("li[id]")%>% html_nodes("h3") %>% 
html_nodes("a[href]") %>%
  html_attr("href") 
college_urls

```


```{r message=FALSE, warning=FALSE}
index_num <- 0

college_tab_1 <-  data.frame("URL" = gsub(" ", "", paste("https://www.usnews.com",college_urls, sep = "")), 
"CollegeName"= "", "TuitionFees" = 0, "RoomBoard" = 0, "TotalEnrollment" = 0, "SchoolType" = "", "YearFounded" = 0, "Setting" = "", "Endowment2017" = 0, "MedianStartingSalaryOfAlumni" = 0, "Selectivity" = "", "Fall2017AcceptanceRate" = 0, "MalePercentage" = 0, "FourYearGraduationRate" = 0, stringsAsFactors = FALSE) 

college_tab_1
```


<<<<<<< HEAD
=======
```{r message=FALSE, warning=FALSE}
#Shawdi start here

get_info <- function(url){
  attr <- url %>% html_node("body") %>% html_nodes("div[id~=content-main]") %>%   
    html_nodes("section[class~=hero-stats-widget-stats]") %>%
    html_nodes("ul") %>% html_nodes("li") %>% html_nodes("strong")
}

get_tuition_rm <- function(info, num){
  a_1 <- info[num] %>%  html_text()
  tuition_rm <- 
    paste(substring(a_1, 2, str_locate(a_1, ",")[1] - 1), substring(a_1, str_locate(a_1, ",")[1] + 1, str_locate(a_1, " ")[1] - 1), sep="")
  tuition_rm
}

get_enrollment <- function(info){
  a_1 <- info[3] %>%  html_text()
  paste(substring(a_1, 1, str_locate(a_1, ",")[1] - 1), substring(a_1, str_locate(a_1, ",")[1] + 1), sep="")
}
```

>>>>>>> 33ffde3edf8bc5185f2b4baab4fbe0bdb751a5af

```{r}
college_tab <- college_tab_1

for (i in 1:nrow(college_tab)){
  url_html <- college_tab[i,1] %>%read_html()
  college_tab[i,]$CollegeName <- url_html %>% html_node("body") %>% html_nodes("h1[class~=hero-heading]") %>% html_text()
<<<<<<< HEAD
  college_tab[i,]$TuitionFees <- get_tuition_rm(get_info(url_html), 1)
  college_tab[i,]$RoomBoard <- get_tuition_rm(get_info(url_html), 2)
  college_tab[i,]$TotalEnrollment <- get_enrollment(get_info(url_html))
  college_tab[i,]$GenderDistribution <- get_gender_ratio(url_html)
  
  #row$Fall2017AcceptanceRate <- url %>% 

=======
#  college_tab[i,]$TuitionFees <- as.double(get_tuition_rm(get_info(url_html), 1))
#  college_tab[i,]$RoomBoard <- as.double(get_tuition_rm(get_info(url_html), 2))
  college_tab[i,]$TotalEnrollment <- as.double(get_enrollment(get_info(url_html)))
>>>>>>> 33ffde3edf8bc5185f2b4baab4fbe0bdb751a5af
  college_tab[i,]$Fall2017AcceptanceRate <- url_html %>% html_node("span[data-test-id~=r_c_accept_rate]") %>% html_text()
  college_tab[i,]$Selectivity <- url_html %>% html_node("span[data-test-id~=c_select_class]") %>% html_text()
  college_tab[i,]$FourYearGraduationRate <- url_html %>% html_node("span[data-test-id~=grad_rate_4_year]") %>% html_text()
}

```


```{r}
#fix College Name formatting
college_tab$CollegeName <- sub("^\n        ","",college_tab$CollegeName)
college_tab$CollegeName <-sub("\n    ","",college_tab$CollegeName)
#fixing Acceptance Rate formatting
college_tab$Fall2017AcceptanceRate <- sub("\n            ","",college_tab$Fall2017AcceptanceRate)
college_tab$Fall2017AcceptanceRate <- sub("%","",college_tab$Fall2017AcceptanceRate)
college_tab$Fall2017AcceptanceRate <- as.double(college_tab$Fall2017AcceptanceRate)
college_tab$Fall2017AcceptanceRate <- college_tab$Fall2017AcceptanceRate/100
#fixing Grad Rate formatting
college_tab$FourYearGraduationRate <- sub("\n            ","",college_tab$FourYearGraduationRate)
college_tab$FourYearGraduationRate <- sub("%","",college_tab$FourYearGraduationRate)
college_tab$FourYearGraduationRate <- as.double(college_tab$FourYearGraduationRate)
college_tab$FourYearGraduationRate <- college_tab$FourYearGraduationRate/100
#fixing Selectivity formatting
#college_tab$Selectivity <- sub("\n            ","",college_tab$Selectivity)
#college_tab$Selectivity <- as.factor(college_tab$Selectivity)

college_tab
```



<<<<<<< HEAD

=======
```{r}
college_tab[i,1] %>%read_html() %>% html_node("span[data-test-id~=c_select_class]") %>% html_text()
```
>>>>>>> 33ffde3edf8bc5185f2b4baab4fbe0bdb751a5af
