---
title: "320 Final Project"
author: "Amy Vuong, Katheryn Fox, Shawdi Hootkany"
date: "5/22/2019"
output: word_document
margin: 1in
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```
#Introduction: 
Going to and succeeding in college involves a lot of different factors, from enrollment size to the cost of tuition. We will explore how these factors are related to other factors, such as graduation rate and the starting salary of gradutes. 

The following dataset contains information about an array of US Universities based on National Ranking. The most important piece of information this dataset provides is the ranking of each university. There are often many assumptions made based on the ranking of a school, such as average starting salary, average tuition, and graduation rate. By having information such as tuition rate, enrollments, location, and median starting salary of alumni, we will be able to test whether there is an actual correlation between ranking and these assumptions. We will also be able to come up with our own predictions and test whether we will be able to predict information such as starting salary based on the provided information about each undergraduate institution. A couple other questions we would like to answer are as follows:
What is the most important factor when predicting starting salary for undergraduate institutions?
Is the cost of undergraduate tuition a key factor?
Does undergraduate salary correlated with the male to female ratio at the institution in any kind of way?
Which variables can be used to predict the cost of tuition?
Which predictors are influential in this?
Can we predict graduation rate?

##Required Tools

You will use R and the following libraries: 
-ggplot2
-rvest
-tidyverse
-stringr

***

#Part 1: Data Scraping

The main url we will be using contains very limited information about each of the schools, such as ranking and tuition, therefore the first step that needs to be taken to be able to gain all the information we need to be able to analyze the data and make predictions is to parse the information into readable data. Detailed information about each school is spread across multiple websites so we will need to retrieve the proper url for each university from the US News website containing the ranking and then parse important information into tables that could be used for data analysis.

We are scraping the data of 100 schools from https://www.usnews.com/best-colleges/rankings/national-universities.
The data we have is stored in a text file since it loads on the page in increments. We parse the data to find the URL for each college's informational page. 

Note: The information for the University of California--Davis was removed from the dataset because it didn’t contain median alumni salary, which plays a large role in our analysis. 

Note: Room and Board, Tuition and Fees, and Median Alumni Salary are all in thousands of dollars.

```{r message=FALSE, warning=FALSE}
library(rvest)
library(tidyverse)
url <-"html_top100.txt"
college_urls <- url %>%
  read_html() %>%
  html_node("body") %>% html_nodes("ol[class~=bEyEue]") %>% html_nodes("li[id]")%>% html_nodes("h3") %>% 
  html_nodes("a[href]") %>%
  html_attr("href") 
head(college_urls)

```

A data frame is created to store the information of each college in rows. Columns are initialized. 

```{r message=FALSE, warning=FALSE}
index_num <- 0

college_tab_1 <-  data.frame("URL" = gsub(" ", "", paste("https://www.usnews.com",college_urls, sep = "")), 
"CollegeName"= "", "TuitionFeesThousands" = 0, "RoomBoardThousands" = 0, "TotalEnrollment" = 0, "SchoolType" = "", "YearFounded" = 0, "Setting" = "", "Endowment2017Millions" = 0, "MedianStartingSalaryOfAlumniThousands" = 0, "Selectivity" = "", "Fall2017AcceptanceRate" = 0, "MalePercentage" = 0, "FourYearGraduationRate" = 0, stringsAsFactors = FALSE) 

#removing one college that doesn't have a median starting salary, for data uniformity
college_tab_1 <- college_tab_1[-c(40),]

head(college_tab_1)
```

Below are functions used to obtain data from the website and parse it.

```{r message=FALSE, warning=FALSE}
#retrieves of vector of size three containing the Tuition&Fees, Room&Board, and total enrollment
get_info <- function(url_html){
  attr <- url_html %>% html_node("body") %>% html_nodes("div[id~=content-main]") %>%   
    html_nodes("section[class~=hero-stats-widget-stats]") %>%
    html_nodes("ul") %>% html_nodes("li") %>% html_nodes("strong")
}

#takes in a vector and index, and parses that information to a double
#ex: $47,263 -> 47263.0
get_tuition_rm <- function(info, num){
  a_1 <- info[num] %>%  html_text()
  tuition_rm <- 
    as.double(paste(substring(a_1, 2, str_locate(a_1, ",")[1] - 1), substring(a_1, str_locate(a_1, ",")[1] + 1, str_locate(a_1, " ")[1] - 1), sep=""))
  tuition_rm / 1000.0
}

#takes in a vector and parses the total enrollment information to a double
get_enrollment <- function(info){
  a_1 <- info[3] %>%  html_text()
  as.double(paste(substring(a_1, 1, str_locate(a_1, ",")[1] - 1), substring(a_1, str_locate(a_1, ",")[1] + 1), sep=""))
}

#gets the percentage of the majority gender at a certain university
get_percent <- function(url_html){
  attr <- url_html %>% html_node("body") %>% html_nodes("div[id~=content-main]") %>%   
    html_nodes("div[class~=block-normal]") %>% html_nodes("span[class~=distribution-breakdown__percentage]") %>% html_text()
  as.double(substring(attr, 1, str_locate(attr, "%")[1] - 1)) / 100.0
}

#retrieves the gender of the majority sex and parses the percentage to be in terms of males
get_gender_ratio <- function(url_html){
  attr <- url_html %>% html_node("body") %>% html_nodes("div[id~=content-main]") %>%   
    html_nodes("div[class~=block-normal]") %>% html_nodes("span[class~=distribution-breakdown__percentage-copy]") %>% html_text()
  attr <- sub("\n                    ","",attr)
  attr <- sub("\n                ","",attr)
  if (attr == "Female"){
    1 - get_percent(url_html)
  }else{
    get_percent(url_html)
  }
}
```

Here, we use both the functions above and the html_node function to fill out the table. 

```{r, message=FALSE, warning = FALSE}
college_tab <- college_tab_1

for (i in 1:nrow(college_tab)){
  url_html <- college_tab[i,1] %>%read_html()
  college_tab[i,]$CollegeName <- url_html %>% html_node("body") %>% html_nodes("h1[class~=hero-heading]") %>% html_text()
  priv_tuition <- url_html %>% html_node("body") %>% html_nodes("span[data-test-id~=v_private_tuition]") %>% html_text()
  college_tab[i,]$TuitionFeesThousands <- ifelse(length(priv_tuition) > 0, priv_tuition, 
                                                 url_html %>% html_node("body") %>% html_node("span[data-test-id~=v_out_state_tuition]") %>% html_text())
  college_tab[i,]$RoomBoardThousands <- url_html %>% html_node("body") %>% html_node("span[data-test-id~=w_room_board]") %>% html_text()
  college_tab[i,]$TotalEnrollment <- url_html %>% html_node("body") %>% html_node("span[data-test-id~=total_all_students]") %>% html_text()
  college_tab[i,]$MalePercentage <- get_gender_ratio(url_html)
  college_tab[i,]$Fall2017AcceptanceRate <- url_html %>% html_node("span[data-test-id~=r_c_accept_rate]") %>% html_text()
  college_tab[i,]$Selectivity <- url_html %>% html_node("span[data-test-id~=c_select_class]") %>% html_text()
  college_tab[i,]$FourYearGraduationRate <- url_html %>% html_node("span[data-test-id~=grad_rate_4_year]") %>% html_text()
  college_tab[i,]$MedianStartingSalaryOfAlumniThousands <-  url_html %>% html_nodes("div[data-field-id=averageStartSalary]") %>%html_node("span[data-test-id]") %>% html_text()
  temp_vector <- url_html %>% html_node("body") %>% html_nodes("div[id~=content-main]") %>%html_nodes("div[class~=flex-row]") %>%   html_nodes("span[class~=heading-small]") %>% html_text()
  college_tab[i,]$SchoolType <- temp_vector[1]
  college_tab[i,]$YearFounded <- temp_vector[2]
  college_tab[i,]$Setting <- temp_vector[5]
  college_tab[i,]$Endowment2017Millions  <- temp_vector[6]
}

head(college_tab)
```


Below, we reformat many of the columns to get usable data. Each column is categorized into the appropriate type of data. 

```{r formatting}
formatted_college_tab <- college_tab
#fix type of School Type, Setting, Year Founded
formatted_college_tab$SchoolType <- as.factor(formatted_college_tab$SchoolType)
formatted_college_tab$Setting <- as.factor(formatted_college_tab$Setting)
formatted_college_tab$YearFounded <- as.integer(formatted_college_tab$YearFounded)
#fix Endowment2017 formatting
formatted_college_tab$Endowment2017Millions  <- ifelse(grepl("billion", formatted_college_tab$Endowment2017Millions ), sub("\\.","",formatted_college_tab$Endowment2017Millions ),formatted_college_tab$Endowment2017Millions )
formatted_college_tab$Endowment2017Millions  <-sub(" billion","00",formatted_college_tab$Endowment2017Millions )
formatted_college_tab$Endowment2017Millions  <-sub(" million","",formatted_college_tab$Endowment2017Millions )
formatted_college_tab$Endowment2017Millions  <-sub("[[:punct:]]", "",formatted_college_tab$Endowment2017Millions )
formatted_college_tab$Endowment2017Millions  <-sub("\\$", "",formatted_college_tab$Endowment2017Millions )
formatted_college_tab$Endowment2017Millions  <-sub(" \\+", "",formatted_college_tab$Endowment2017Millions )
formatted_college_tab$Endowment2017Millions <- as.double(formatted_college_tab$Endowment2017Millions)
#fix College Name formatting
formatted_college_tab$CollegeName <- sub("^\n        ","",formatted_college_tab$CollegeName)
formatted_college_tab$CollegeName <-sub("\n    ","",formatted_college_tab$CollegeName)
#fixing Acceptance Rate formatting
formatted_college_tab$Fall2017AcceptanceRate <- sub("\n            ","",formatted_college_tab$Fall2017AcceptanceRate)
formatted_college_tab$Fall2017AcceptanceRate <- sub("%","",formatted_college_tab$Fall2017AcceptanceRate)
formatted_college_tab$Fall2017AcceptanceRate <- as.double(formatted_college_tab$Fall2017AcceptanceRate)
formatted_college_tab$Fall2017AcceptanceRate <- formatted_college_tab$Fall2017AcceptanceRate/100
#fixing Grad Rate formatting
formatted_college_tab$FourYearGraduationRate <- sub("\n            ","",formatted_college_tab$FourYearGraduationRate)
formatted_college_tab$FourYearGraduationRate <- sub("%","",formatted_college_tab$FourYearGraduationRate)
formatted_college_tab$FourYearGraduationRate <- as.double(formatted_college_tab$FourYearGraduationRate)
formatted_college_tab$FourYearGraduationRate <- formatted_college_tab$FourYearGraduationRate/100
#fixing Salary formatting
formatted_college_tab$MedianStartingSalaryOfAlumniThousands <- 
  sub("\n            ","",formatted_college_tab$MedianStartingSalaryOfAlumniThousands)
formatted_college_tab$MedianStartingSalaryOfAlumniThousands <- gsub("\\*","",formatted_college_tab$MedianStartingSalaryOfAlumniThousands)
formatted_college_tab$MedianStartingSalaryOfAlumniThousands <- gsub("\\$","",formatted_college_tab$MedianStartingSalaryOfAlumniThousands)
formatted_college_tab$MedianStartingSalaryOfAlumniThousands <- gsub("\\,","",formatted_college_tab$MedianStartingSalaryOfAlumniThousands)
formatted_college_tab$MedianStartingSalaryOfAlumniThousands <- as.double(formatted_college_tab$MedianStartingSalaryOfAlumniThousands)/1000
#fixing Selectivity formatting
formatted_college_tab$Selectivity <- sub("\n            ","",formatted_college_tab$Selectivity)
formatted_college_tab$Selectivity <- as.factor(formatted_college_tab$Selectivity)
#fixing Tuition formatting
formatted_college_tab$TuitionFeesThousands <- sub("\n            ", "",formatted_college_tab$TuitionFeesThousands )
formatted_college_tab$TuitionFeesThousands <- sub(" \\(2018-19\\)", "",formatted_college_tab$TuitionFeesThousands )
formatted_college_tab$TuitionFeesThousands  <-sub("\\,", "",formatted_college_tab$TuitionFeesThousands )
formatted_college_tab$TuitionFeesThousands  <-sub("\\$", "",formatted_college_tab$TuitionFeesThousands )
formatted_college_tab$TuitionFeesThousands <- as.double(formatted_college_tab$TuitionFeesThousands)/1000
#fixing RoomBoard formatting
formatted_college_tab$RoomBoardThousands <- sub("\n            ", "",formatted_college_tab$RoomBoardThousands )
formatted_college_tab$RoomBoardThousands <- sub(" \\(2018-19\\)", "",formatted_college_tab$RoomBoardThousands )
formatted_college_tab$RoomBoardThousands  <-sub("\\,", "",formatted_college_tab$RoomBoardThousands )
formatted_college_tab$RoomBoardThousands  <-sub("\\$", "",formatted_college_tab$RoomBoardThousands )
formatted_college_tab$RoomBoardThousands <- as.double(formatted_college_tab$RoomBoardThousands)/1000
#fixing Enrollment formatting
formatted_college_tab$TotalEnrollment <- sub("\n            ", "",formatted_college_tab$TotalEnrollment )
formatted_college_tab$TotalEnrollment  <-sub("\\,", "",formatted_college_tab$TotalEnrollment )
formatted_college_tab$TotalEnrollment <- as.double(formatted_college_tab$TotalEnrollment)


formatted_college_tab <- formatted_college_tab %>% mutate(TotalCostThousands =TuitionFeesThousands + RoomBoardThousands )

formatted_college_tab <- na.omit(formatted_college_tab)
nrow(formatted_college_tab)
as.tibble(formatted_college_tab)
```


```{r}
#to save as csv to easily work on it without having to reload
write.csv(formatted_college_tab, file = "college_info.csv")
```
```{r}
formatted_college_tab <- read.csv("college_info.csv")
formatted_college_tab <- formatted_college_tab[,-c(1)]
as.tibble(formatted_college_tab)
```

***

#Part 2: Data Visualization
We plot the data in order to visualize relationships among the attributes. 
```{r fig1, fig.height = 10, fig.width = 12}
#Starting Salary
#-histograms
library(ggplot2)
plot_1 <- formatted_college_tab %>%
  ggplot(aes(MedianStartingSalaryOfAlumniThousands)) +
    geom_histogram()+ 
    labs(title="Starting Salary Distribution", x="Median Starting Salary of Alumni (Thousands)", y="Count")
plot_1

```

The distribution of the median starting salary of alumni from all the school seems to be a bell-shaped curve (a little skewed right), centering around $55,000.

```{r fig2, fig.height = 10, fig.width = 12}

#Tuition Cost
#-histograms
library(ggplot2)
plot_2 <- formatted_college_tab %>%
  ggplot(aes(TuitionFeesThousands)) +
    geom_histogram()+ 
        labs(title="Tuition Cost Distribution", x="Tuition Cost (Thousands)", y="Count")
plot_2

```

The distribution of tution costs of all the schools is skewed left, with a range of $60,000. 

```{r fig3, fig.height = 10, fig.width = 12}
#Acceptance rate vs graduation rate

library(ggplot2)
plot_3 <- formatted_college_tab %>%
  ggplot(aes(x=Fall2017AcceptanceRate, y=FourYearGraduationRate)) +
    geom_point()+ 
    geom_smooth(method=lm)+
        labs(title="Acceptance Vs. Graduation Rate", x="Fall 2017 Acceptance Rate", y="Four Year Graduation Rate")
plot_3
```
There is a linear relationship between acceptance rate (Fall 2017) and the four year graduation rate. It is an overall negative relationship. The higher the acceptance rate, the lower the rate of graduation. 


```{r fig4, fig.height = 10, fig.width = 12}
#Boxplots of (1) gradruattion rate & (2) admission rate by selectivity 
library(ggplot2)
formatted_college_tab$Selectivity <- factor(formatted_college_tab$Selectivity, c("Selective","More selective","Most selective"))
plot_4 <-  formatted_college_tab %>%
  ggplot(aes(x=Selectivity, y=FourYearGraduationRate)) +
    geom_boxplot()+
        labs(title="Graduation Rate based on Selectivity", x="Selectivity Level", y="Four Year Graduation Rate")
plot_4

```
This is significant difference in four year graduation rates based on their Selectivity Level of accepting students. These boxplots show that each 3 selectivity level vary significantly on range and central tendency. The more selective a college is, the greater their graduation rates seem to be. 

```{r fig5, fig.height = 10, fig.width = 12}
#Setting vs. room board

library(ggplot2)
formatted_college_tab$Setting <- factor(formatted_college_tab$Setting, c("Rural","Suburban","Urban", "City"))
plot_5 <- formatted_college_tab %>%
  ggplot(aes(x=Setting, y=RoomBoardThousands)) +
    geom_boxplot()+
        labs(title="Setting vs. Room & Board Costs", x="Setting", y="Room & Board Costs (Thousands)")
plot_5

```
The boxplots of room & board costs based on setting shows that the setting of the college has some influence the room and board costs for the students. The median room and board costs of the City settingvary from that of the others. The spread is also greater for the City setting while it is much smaller for the rural setting. 

```{r fig6, fig.height = 10, fig.width = 12}
plot_6 <- formatted_college_tab %>%
  
  ggplot(aes(x=TotalCostThousands, y=MedianStartingSalaryOfAlumniThousands)) +
    geom_point()+ 
    geom_smooth(method=lm)+
        labs(title="Total Cost vs. Median Starting Salary", x="Total Cost (Thousand)", y="Median Starting Salary Of Alumni (Thousands)")
plot_6
```

There appears to be a positive linear relationship between median starting salary and total cost of colleges. The general trends shows that the more students spend on tution, room, and board, the more likely that their starting salary is higher. 

```{r fig7, fig.height = 10, fig.width = 12}
plot_7 <- formatted_college_tab %>%
  ggplot(aes(x=SchoolType, y=MedianStartingSalaryOfAlumniThousands  
)) +
    geom_boxplot()+
        labs(title="Median Starting Salary Of Alumni Based on School Type  ", x="School Type", y="Median Starting Salary Of Alumni (Thousands)")
plot_7
```
Between school types, private colleges seem to have greater starting salaries than public schools, based on the medians of these boxplots. 
```{r}
 formatted_college_tab %>% group_by(Selectivity) %>%
  summarise(n())
```

```{r fig8, fig.height = 10, fig.width = 12}
plot_8 <- formatted_college_tab %>%
  ggplot(aes(x=MalePercentage, y=MedianStartingSalaryOfAlumniThousands  
)) +
    geom_point()+
  geom_smooth(method=lm)+
        labs(title="Male Percentage vs. Median Starting Salary of Alumni  ", x="Male Percentage", y="Median Starting Salary Of Alumni (Thousands)")
plot_8
```
Although the points are scattered with some variation, there is a general positive correlation between median starting salary of alumni and the male percentage of the student body of colleges.

***

#Part 3: Model Fitting and Selection

##Fitting model for tuition prices

```{r}
#adjusting dataset to remove variables not able to be used in model fitting
college_info <- formatted_college_tab[,-c(1,2)]
head(college_info)
```

```{r}
college_info$FourYearGraduationRate <- college_info$FourYearGraduationRate*100
college_info$MalePercentage <- college_info$MalePercentage*100
college_info$Fall2017AcceptanceRate <- college_info$Fall2017AcceptanceRate*100
```

```{r}
#linear model fitting 
tuition_lm_1 <- lm(TuitionFeesThousands~.-RoomBoardThousands-TotalCostThousands, data = college_info)
summary(tuition_lm_1)
```

```{r}
plot(tuition_lm_1)
```

```{r}
tuition_lm_2 <- step(tuition_lm_1, direction = "both", steps = 1000, trace = F)
summary(tuition_lm_2)
```

```{r}
plot(tuition_lm_2)
```
```{r}
anova(tuition_lm_2,tuition_lm_1, test="Chisq")
```

##Fitting model for graduation rate
```{r}
#linear model fitting 
gradrate_lm_1 <- lm(MedianStartingSalaryOfAlumniThousands~.-TuitionFeesThousands-RoomBoardThousands, data = na.omit(college_info))
summary(gradrate_lm_1)
plot(gradrate_lm_1)
```
```{r}
gradrate_lm_2 <- step(gradrate_lm_1, direction = "both", steps = 1000, trace = F)
summary(gradrate_lm_2)

```
```{r}
plot(gradrate_lm_2)

anova(gradrate_lm_2,gradrate_lm_1, test="Chisq")
```



***

#Conclusion

Being aware of all these factors in succeeding in college is very important when deciding where to go. 

References: 
-College Ranking Data: https://www.usnews.com/best-colleges/rankings/national-universities

